\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-11-09 doc/USER_MANUAL/05_running_the_solver.tex
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

\noindent
Most parameters in the \texttt{Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.\\

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\begin{centering}
\includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.jpg}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

@@

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
function (i.e., the second derivative of a Gaussian) when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
in the main parameter file {\texttt{Par\_file}}. In case that the
solver uses a (pseudo) Dirac delta source time function to represent
a force point source, a very short duration of five time steps is automatically set by default.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the force
source in the East/North/Vertical basis (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time
shift}. Each subevent can have its own half latitude, longitude, depth,
\texttt{half duration} and force parameters.

@@

@@

In addition to inbuild source time function, the solver can also use an external source time function defined by the user. This option can be
activated by setting   \texttt{EXTERNAL\_SOURCE\_FILE} to true in \texttt{Par\_File} and by adding the name of the file
containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
The source time function file must contain the value of the time step on its first line, and then a single column with the amplitude of the source time function for all the time steps.
The time step must be exactly the same as that used for the simulation.
Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in
the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error.
When using an external source file, you can still set up the source location and directivity as in the default case.
In the FORCESOLUTION file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source. Then if you want to define a directivity, then change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP". What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source time function from an external file.

The solver can calculate seismograms at any number of stations for
basically the same numerical cost, so the user is encouraged to include
as many stations as conceivably useful in the \texttt{STATIONS} file,
which looks like this:
%
\begin{figure}[H]
\begin{centering}
\includegraphics[width=4in]{figures/STATIONS_basin_explained.jpg}
\par
\end{centering}
\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
be provided in geographical coordinates. The width of the station
label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
in the \texttt{constants.h} file), and the network label should be
no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
in the \texttt{constants.h} file).}
\label{fig:Sample-STATIONS-file.}
\end{figure}

\noindent
Each line represents one station in the following format:
{\small
\begin{verbatim}
Station Network Latitude(degrees) Longitude(degrees) Elevation(m) burial(m)
\end{verbatim}
}
The solver \texttt{xspecfem3D} filters the list of stations in file
\texttt{DATA/STATIONS} to exclude stations that are not located within
the region given in the \texttt{Par\_file} (between \texttt{LATITUDE\_MIN}
and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
Elevation and burial are generally applicable to geographical regions.
Burial is measured down from the top surface.\\

@@

\item {\it Procedure 2:} useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface
\begin{enumerate}
\item [-] In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the Par\_file, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the Par\_file is on, this remark applies to the position of the receivers as well.\\
\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should then look like:
{\small \begin{verbatim}
BONE  GR  y1  x1  0.00  z1
BONE  GR  y2  x2  0.00  z2
\end{verbatim}}
The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
burial as the $z$ coordinate.
Third column is Y and Fourth is X due to the latitude/longitude convention.\\
\end{enumerate}

@@

Solver output is provided in the \texttt{OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{constants.h}:
\begin{verbatim}
! uncomment this to write messages to the screen
! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
\end{verbatim}
On PC clusters the seismogram files are generally written to the local
disks (the path \texttt{LOCAL\_PATH} in the \texttt{Par\_file}) and
need to be gathered at the end of the simulation.

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}' files in the \texttt{\small OUTPUT\_FILES/} directory. These tiny
files look something like this:
{\small
\begin{verbatim}
Time step #          10000
Time:     108.4890      seconds
Elapsed time in seconds =    1153.28696703911
Elapsed time in hh:mm:ss =     0 h 19 m 13 s
Mean elapsed time per time step in seconds =     0.115328696703911
Max norm displacement vector U in all slices (m) =     1.0789589E-02
\end{verbatim}
}
The \texttt{\small timestamp{*}} files provide the
\texttt{\small Mean elapsed time per time step in seconds}, which may be used
to assess performance on various machines (assuming you are the only
user on a node), as well as the
\texttt{\small Max norm displacement vector U in all slices~(m)}.
If something is wrong with the
model, the mesh, or the source, you will see the code become unstable
through exponentially growing values of the displacement and fluid
potential with time, and ultimately the run will be terminated by
the program. You can control the rate at which the timestamp files
are written based upon the parameter
\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}
in the \texttt{\small Par\_file}.

Having set the \texttt{Par\_file} parameters, and having provided
the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
files, you are now ready to launch the solver! This is most easily
accomplished based upon the \texttt{go\_solver} script (See Chapter
\ref{cha:Scheduler} for information about running through a scheduler,
e.g., LSF). You may need to edit the last command at the end of the
script that invokes the \texttt{mpirun} command. The \texttt{runall}
script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
in sequence. This is a safe approach that ensures using the correct
combination of distributed database output and solver input.

@@

\noindent
Regarding attenuation (viscoelasticity), in the Par\_file you need to select the number of standard linear solids (N\_SLS) to use to mimic a constant $Q$ quality factor.
Using N\_SLS = 3 is always safe. If (and only if) you know what you are doing, you can try to reduce that in order to reduce the cost of the simulations.
Figure~\ref{fig:selectNSLS} shows values that you can consider using (again, if and only if you know what you are doing). That table has been created by Zhinan Xie using
a comparison between results obtained with a truly-constant $Q$ and results obtained with its approximation based on N\_SLS standard linear solids.
The comparison is performed using the time-frequency misfit and goodness-of-fit criteria proposed by \cite{Kristekova_2009}.
The table is drawn for a dimensionless parameter representing the distance of propagation.
%%
\begin{figure}[htbp]
\centering
\includegraphics[width=5in]{figures/minimum_number_of_SLS_that_can_be_used_in_viscoelastic_simulation.png}
\caption{Table showing how you can select a value of N\_SLS smaller than 3, if and only if you know what you are doing.}
\label{fig:selectNSLS}
\end{figure}
%%

>>>>

\noindent
Most parameters in the \texttt{DATA/Par\_file} should be set prior to running
the databases generation. Only the following parameters may be changed
after running \texttt{xgenerate\_databases}:
\begin{itemize}
\item the simulation type control parameters \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{PML\_CONDITIONS}
on condition that the\\
 \texttt{PML\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified
after running the databases generation.
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered{} option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{MOVIE\_TYPE}, \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO} and\\
\texttt{NTSTEP\_BETWEEN\_OUTPUT\_SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{DATA/Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver
\texttt{xspecfem3D}.\\

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\begin{centering}
\includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.jpg}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

@@

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
function (i.e., the second derivative of a Gaussian) when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
in the main parameter file \texttt{DATA/Par\_file}. In case that the
solver uses a (pseudo) Dirac delta source time function to represent
a force point source, a very short duration of five time steps is automatically set by default.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the force
source in the East/North/Vertical basis (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time
shift}. Each subevent can have its own latitude, longitude, depth,
\texttt{half duration} and force parameters.

@@

In addition to inbuild source time function, the solver can also use an external source time function defined by the user. This option can be activated by setting \texttt{USE\_EXTERNAL\_SOURCE\_FILE} to true in \texttt{DATA/Par\_file} and by adding the name of the file containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
The source time function file must contain the value of the time step on its first line, and then a single column with the amplitude of the source time function for all the time steps. The time step must be exactly the same as that used for the simulation. Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error. When using an external source file, you can still set up the source location and directivity as in the default case.
In the FORCESOLUTION file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source. Then if you want to define a directivity, then change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP". What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source time function from an external file.

@@

The solver can calculate seismograms at any number of stations for
basically the same numerical cost, so the user is encouraged to include
as many stations as conceivably useful in the \texttt{STATIONS} file,
which looks like this:
%
\begin{figure}[H]
\begin{centering}
\includegraphics[width=4in]{figures/STATIONS_basin_explained.jpg}
\par
\end{centering}
\caption{Sample \texttt{STATIONS} file. Station latitude and longitude should
be provided in geographical coordinates. The width of the station
label should be no more than 32 characters (see \texttt{MAX\_LENGTH\_STATION\_NAME}
in the \texttt{setup/constants.h} file), and the network label should be
no more than 8 characters (see \texttt{MAX\_LENGTH\_NETWORK\_NAME}
in the \texttt{setup/constants.h} file).}
\label{fig:Sample-STATIONS-file.}
\end{figure}

\noindent
Each line represents one station in the following format:
{\small
\begin{verbatim}
Station Network Latitude(degrees) Longitude(degrees) Elevation(m) burial(m)
\end{verbatim}
}
The solver \texttt{xspecfem3D} filters the list of stations in file
\texttt{DATA/STATIONS} to exclude stations that are not located within
the region given in the \texttt{Mesh\_Par\_file} (between \texttt{LATITUDE\_MIN}
and \texttt{LATITUDE\_MAX} and between \texttt{LONGITUDE\_MIN} and
\texttt{LONGITUDE\_MAX}). The filtered file is called \texttt{DATA/STATIONS\_FILTERED}.
Elevation and burial are generally applicable to geographical regions.
Burial is measured down from the top surface.\\

@@

\item {\it Procedure 2:} useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface
\begin{enumerate}
\item [-] In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\\
\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should then look like:
{\small \begin{verbatim}
BONE  GR  y1  x1  0.00  z1
BONE  GR  y2  x2  0.00  z2
\end{verbatim}}
The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
burial as the $z$ coordinate.
Third column is Y and Fourth is X due to the latitude/longitude convention.\\
\end{enumerate}

@@

Solver output is provided in the \texttt{OUTPUT\_FILES} directory
in the \texttt{output\_solver.txt} file. Output can be directed to
the screen instead by uncommenting a line in \texttt{setup/constants.h}:
\begin{verbatim}
! uncomment this to write messages to the screen
! integer, parameter :: IMAIN = ISTANDARD_OUTPUT
\end{verbatim}
On PC clusters the seismogram files are generally written to the local
disks (the path \texttt{LOCAL\_PATH} in the \texttt{DATA/Par\_file} and
need to be gathered at the end of the simulation.

While the solver is running, its progress may be tracked by monitoring
the `\texttt{\small timestamp{*}}' files in the \texttt{\small OUTPUT\_FILES/} directory. These tiny
files look something like this:
{\small
\begin{verbatim}
Time step #          10000
Time:     108.4890      seconds
Elapsed time in seconds =    1153.28696703911
Elapsed time in hh:mm:ss =     0 h 19 m 13 s
Mean elapsed time per time step in seconds =     0.115328696703911
Max norm displacement vector U in all slices (m) =     1.0789589E-02
\end{verbatim}
}
The \texttt{\small timestamp{*}} files provide the
\texttt{\small Mean elapsed time per time step in seconds}, which may be used
to assess performance on various machines (assuming you are the only
user on a node), as well as the
\texttt{\small Max norm displacement vector U in all slices~(m)}.
If something is wrong with the
model, the mesh, or the source, you will see the code become unstable
through exponentially growing values of the displacement and fluid
potential with time, and ultimately the run will be terminated by
the program. You can control the rate at which the timestamp files
are written based upon the parameter
\texttt{\small NTSTEP\_BETWEEN\_OUTPUT\_INFO}
in the \texttt{DATA/Par\_file}.

Having set the \texttt{DATA/Par\_file} parameters, and having provided
the \texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and \texttt{STATIONS}
files, you are now ready to launch the solver! This is most easily
accomplished based upon the \texttt{go\_solver} script (See Chapter
\ref{cha:Scheduler} for information about running through a scheduler,
e.g., LSF). You may need to edit the last command at the end of the
script that invokes the \texttt{mpirun} command. The \texttt{runall}
script compiles and runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D}
in sequence. This is a safe approach that ensures using the correct
combination of distributed database output and solver input.

@@

\noindent
Regarding attenuation (viscoelasticity), in the \texttt{setup/constants.h} you need to select the number of standard linear solids (N\_SLS) to use to mimic a constant $Q$ quality factor.
Using N\_SLS = 3 is always safe. If (and only if) you know what you are doing, you can try to reduce that in order to reduce the cost of the simulations.
Figure~\ref{fig:selectNSLS} shows values that you can consider using (again, if and only if you know what you are doing). That table has been created by Zhinan Xie using
a comparison between results obtained with a truly-constant $Q$ and results obtained with its approximation based on N\_SLS standard linear solids.
The comparison is performed using the time-frequency misfit and goodness-of-fit criteria proposed by \cite{Kristekova_2009}.
The table is drawn for a dimensionless parameter representing the distance of propagation.
%%
\begin{figure}[htbp]
\centering
\includegraphics[width=5in]{figures/minimum_number_of_SLS_that_can_be_used_in_viscoelastic_simulation.png}
\caption{Table showing how you can select a value of N\_SLS smaller than 3, if and only if you know what you are doing.}
\label{fig:selectNSLS}
\end{figure}
%%

### 2022-11-09 doc/USER_MANUAL/06_fault_sources.tex
Buildbot ran the code beautifier.
Remove a conflict
Path updates for parameters moved from constants.h to Par_file and vice versa, and other minor corrections.

The fault-id identifies the fault that contains the station. It is
the index of appearance in the faults list after line 2 of Par\_file\_faults

@@

To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set TPV16=.TRUE..

\end{lyxlist}
\begin{lyxlist}{00.00.0000}
\item [{\textbf{DATA/rsf\_hete\_input\_file.txt}}]  Heterogeneous stresses and friction input for
  rate and state friction.
To activate this feature, in Par\_file\_faults name list \&\textbf{RUPTURE\_SWITCHES}, set RSF\_HETE=.TRUE..
  The format of  {\textbf{DATA/rsf\_hete\_input\_file.txt}} is as such, in the first line there are four integers
  that are sequentially documenting the number of divisions along strike, number of divisions along
  dip, cell size along strike , cell size along dip. Then the following N (N=NumberOfDivisionsAlongStrike
  * NumberOfDivisionsAlongDip) lines will document the stress and friction properties on the grid.
  There are a total of 13 columns.
  \begin{enumerate}
  \item{Column1} = Along strike distance(m)

@@

The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
is to damp spurious oscillations generated by the fault slip at frequencies
that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
(in seconds) depends on the size of the elements on the fault. Here
is how to set it:
\begin{enumerate}
\item Determine the average linear size of the elements on the fault plane,
$\mathtt{h\_fault}$. Usually this value is prescribed by the user
during mesh generation. Otherwise it can be found by inspection of
the mesh inside the CUBIT GUI.
\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
compute\\
 $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
\\
This is the critical time step in an elastic medium for a hypothetical
element of cubic shape with size equal to $\mathtt{h\_fault}$.
\item Set $\mathtt{eta}$ in $\mathtt{Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
A larger $\mathtt{eta}$ damps high-frequencies more aggressively
but it might also affect lower frequencies and rupture speed.
\end{enumerate}
Viscosity reduces numerical stability: the critical timestep in a
simulation with Kelvin-Voigt damping needs to be smaller than that
in a purely elastic simulation. Here is how to set the time step accordingly:
\begin{enumerate}
\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
a few time steps)
\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
\item Reset the timestep of the simulation with a Kelvin-Voigt material
to a value smaller than\\
 $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
\end{enumerate}
Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
because elements off the fault might be smaller or more distorted
than element faces on the fault.

>>>>

The fault-id identifies the fault that contains the station. It is
the index of appearance in the faults list after line 2 of \texttt{DATA/Par\_file\_faults}

@@

To activate this feature, in \texttt{DATA/Par\_file\_faults} name list \&\textbf{RUPTURE\_SWITCHES}, set TPV16=.TRUE..

\end{lyxlist}
\begin{lyxlist}{00.00.0000}
\item [{\textbf{DATA/rsf\_hete\_input\_file.txt}}]  Heterogeneous stresses and friction input for
  rate and state friction.
To activate this feature, in \texttt{DATA/Par\_file\_faults} name list \&\textbf{RUPTURE\_SWITCHES}, set RSF\_HETE=.TRUE..
  The format of  {\textbf{DATA/rsf\_hete\_input\_file.txt}} is as such, in the first line there are four integers
  that are sequentially documenting the number of divisions along strike, number of divisions along
  dip, cell size along strike , cell size along dip. Then the following N (N=NumberOfDivisionsAlongStrike
  * NumberOfDivisionsAlongDip) lines will document the stress and friction properties on the grid.
  There are a total of 13 columns.
  \begin{enumerate}
  \item{Column1} = Along strike distance(m)

@@

The purpose of the Kelvin-Voigt viscosity in the dynamic fault solver
is to damp spurious oscillations generated by the fault slip at frequencies
that are too high to be resolved by the mesh. The viscosity $\mathtt{eta}$
(in seconds) depends on the size of the elements on the fault. Here
is how to set it:
\begin{enumerate}
\item Determine the average linear size of the elements on the fault plane,
$\mathtt{h\_fault}$. Usually this value is prescribed by the user
during mesh generation. Otherwise it can be found by inspection of
the mesh inside the CUBIT GUI.
\item Use the Matlab function $\mathtt{utils/critical\_timestep.m}$ to
compute\\
 $\mathtt{dtc\_fault}=\mathtt{critical\_timestep\left(c_{p},h\_fault,ngll\right)}$.
\\
This is the critical time step in an elastic medium for a hypothetical
element of cubic shape with size equal to $\mathtt{h\_fault}$.
\item Set $\mathtt{eta}$ in $\mathtt{DATA/Par\_file\_faults}$ to (0.1 to 0.3)$\ensuremath{\times}\mathtt{dtc\_fault}$.
A larger $\mathtt{eta}$ damps high-frequencies more aggressively
but it might also affect lower frequencies and rupture speed.
\end{enumerate}
Viscosity reduces numerical stability: the critical timestep in a
simulation with Kelvin-Voigt damping needs to be smaller than that
in a purely elastic simulation. Here is how to set the time step accordingly:
\begin{enumerate}
\item Run a test simulation without viscosity ($\mathtt{eta}$=0 and only
a few time steps)
\item Look for the \textquotedbl{}maximum suggested time step\textquotedbl{}
in \texttt{OUTPUT\_FILES/output\_mesher.txt}. This is the critical
timestep of a purely elastic simulation, $\mathtt{dtc\_bulk}$.
\item Reset the timestep of the simulation with a Kelvin-Voigt material
to a value smaller than\\
 $\mathtt{dtc\_kv}=\mathtt{eta}\left(\sqrt{1+\mathtt{dtc\_bulk^{2}}/\mathtt{eta^{2}}}-1\right)$
\end{enumerate}
Note that in general $\mathtt{dtc\_bulk}$ is smaller than $\mathtt{dtc\_fault}$,
because elements off the fault might be smaller or more distorted
than element faces on the fault.
